### 2009-06-14 doc/tech_doc/infosys/infosys_technical.tex
Temporary changes for the internal review

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%\usepackage{trackchanges}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2CM}
\hypersetup{
  pdfauthor = {},
  pdftitle = {Documentation of the Web Service based ARC Information System},
  pdfsubject = {},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large \footnote{@} \large \par}
\end{titlepage}

@@

\subsection{Functionality}
\label{sub:peer_to_peer_functionality}
The ISIS nodes are able to organize themselves into a peer-to-peer network by synchronizing the stored replicated database and maintain this synchronicity and the network topology. Every node of this peer-to-peer cloud (that is an ISIS service) has an own PeerID, an identifier only used in this network, made as a hash of its endpoint URL. The members are ordered by this identifier where there is also a sequence defined between them where the successor entity is the ISIS with the smallest PeerID of the greater PeerIDs. This metric is used for defining the neighbour relation that is the base of the inter-peer-to-peer network communication. It is presumed that only registered ISIS nodes can have any role in the cloud.

\subsubsection{Connecting to the peer-to-peer network}
Every ISIS tries to connect to the network at its startup by accessing the InfoProvider ISIS nodes known from its configuration. The choice from among the few pre-configured InfoProviders is random to achieve as much load-balancing as possible. (In the design it is assumed that a huge amount of ISIS nodes are configured in very similar way, where the InfoProvider ISIS nodes are mainly the same.) If there is no InfoProvider available or nothing is known from the configuration, the node decides to be the fist member and to build a new network.

If there are any InfoProvider nodes available the list of member ISIS nodes will be queried and the ÒsuccessorÓ will be defined using the previously mentioned metric. This ÒsuccessorÓ node will help the entering ISIS to connect to the cloud. The new member will use ÒConnectÓ Web Service operation to get the newest version of database stored in the peer-to-peer network, and place its own entries to the cloud by registering them. An initial database synchronization can be achieved using this method.

We can prevent the InfoProvider ISIS nodes from overloading by balancing the operation of database synchronizing being the most costly procedure in the connection phase.

There is also an other extension used regarding the InfoProvider ISIS nodes. If a node is not able to connect any of the InfoProviders then it tries to repeat it later. This results that the network is able to fuse two disjoint parts of network when the missing InfoProvider ISIS appears again.

\subsubsection{Routing in the peer-to-peer network}
We are keeping the databases in synch by passing every Registration and RemoveRegistrations to every node acting in the network. This yields redundancy and the fault tolerant message routing is very important between any two ISIS nodes.

The most important parameter of the network is the {\it sparsity}. This is an integer not less then 2. It determines the number of neighbors in the function of the actual number of member nodes of the network, where the number of neighbors can change when new members connecting to the network or members leave. A sparse graph can be shaped by choosing a great value for sparsity or a dense one by choosing the less. This value is defined during the design and determines the main properties of the peer-to-peer network.
\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{Circle2-1.png}}}
\caption{\label{fig:p2p_neighbors_2} The neighbor connections in the peer-to-peer network in case of sparsity=2} }
\end{figure}

@@

The registration of service is carried out by internal module called Registrant. The Registrant is active module of the HED (Hosting Environment Daemon) which is bound to a set of ISISes. In practice, the configuration part of the Registrant contains exactly one ISIS to bind, and the Registrant will collect the necessary information about the other ISISes belonging to the same network.

To register services to more than one ISIS network multiple Registrant instances has to be configured. In this case, the default Registrant will be used for registering every services unless configured explicitly.
The registration of service can be done once or periodically based either on the configuration of the Registrant or overwritten for every service separately. The Registrant is also performing message aggregation of all services linked to it if possible. The simplified algorithm of the Registrant is presented below.

\begin{framed}
  Registrant - pseudo algorithm\\
  \\
  \verb#// Initialize phase#\\
  \verb#Read the configuration and store the information about the services in a list#\\
  \verb#do { // Cyclic phase in a different Thread#\\
  \verb# wake_up_time = now();#\\
  \verb# messages = null;#\\
  \verb#  if ( 0 < count(service where service.next_run <= wake_up_time)) {#\\
  \verb#    foreach( service where service.next_run <= wake_up_time) {#\\
  \verb#      messages.add(service.RegistrationCollector);#\\
  \verb#      service.next_run = wake_up_time + service.period;#\\
  \verb#    }#\\
  \verb#    if (0 < count(messages)) {#\\
  \verb#      sent_message = assemble message with headers(messages);#\\
  \verb#      send(sent_message);#\\
  \verb#    }#\\
  \verb#  } else {#\\
  \verb#    sleep(min(service.next_run) - now()); #\\
  \verb#  }#\\
  \verb#} while(true)#\\
\end{framed}
Current implementation does not allow value of the period to be less than 2 minutes.

Service provides \textbf{Service Advertisement} part of information sent to ISIS (see Section \ref{sub:isis_interface}).
Before sending this information the Registrant extends it with additional data (\textbf{Service Advertisement Metadata}).

An example layout of services is shown on Figure \ref{fig:registration_overview}. In this configuration example 
there are two Registrants configured in one HED container for three services. The \textit{Registrant A} is the 
default Registrant and the services configured in the following way:

\begin{itemize}
  \item Service 1: There is no Registrant configured so the default one will register it.
  \item Service 2: The \textit{Registrant A} is configured explicitly.
  \item Service 3: The \textit{Registrant B} is configured explicitly.
\end{itemize}

So the Service 1 and 2 are handled both by \textit{Registrant A} and Service 3 by \textit{Registrant B}.
In the first step each Registrant performs information collection from all assigned services sequentially. Registrants themselves are executing in parallel. For Figure \ref{fig:registration_overview} the approximate sequence of information collection is (\textit{A1}-\textit{A2}, \textit{B1}-\textit{B2}) followed by (\textit{A3}, \textit{A4}). After collection  the aggregated information is sent to the corresponding ISISes independently by each other Registrant (\textit{A5}, \textit{B5}).

@@

For each information provided by service through the LIDI interface same procedure as described in 
Section \ref{sub:isis_authorization_information_authorization} should be applied.

@@

The registration operation is done by the Registrants as described in Section \ref{sec:service_registration}.
Each Registrant is instantiated by corresponding InfoRegistrar XML element in configuration file as 
defined in Section \ref{annex:service_configuration_schema}. First InfoRegistrar defines Registrant 
which is used to register every service by default.

Registration parameters per service are defined by InfoRegister element located inside corresponding 
service configuration element. To assign service to specific Registrant(s) one may use Registrar 
configuration elements situated inside InfoRegister. If none Registrar element is defined service will 
be registered using first (default) Registrant. To make not register itself special configuration element 
NoRegister has to be used.

Following elements are defined inside configuration element of the Registrant:
\begin{description}
\item{URL} specifies contact endpoint of a bootstrap ISIS. If neede further ISIS addresses will be 
queried from this service. This element is mandatory.
\item{KeyPath, CertificatePath, ProxyPath and CACertificatesDir} are paths to files storing X509 
credentials used for establishing connections. Those elements are optional and needed only if 
TLS communication is used.
\item{Retry} specifies how many times communication to ISIS have to fail/timeout to start treating is
as unavailable. The default value is 5.
\end{description}

@@

    <xsd:complexType name="InfoRegistrar_Type">
        <!-- This element defines configuration of Information Registration
             active element. This element is located at top level configuration
             document. Each instance if such element instatiates one instance
             of Information Registrant active element. -->
        <xsd:sequence>
            <!-- URL specifies contact endpoint of a bootstrap Information
                 Registration service. Further ISIS adresses will be queried
                 from this service. -->
            <xsd:element name="URL" type="xsd:string"/> <!-- maxOccurs="unbounded" -->
            <!-- Optional KeyPath, CertificatePath, ProxyPath and CACertificatesDir are
                 paths to files storing X509 credentials used for establishing connections. -->
            <xsd:element name="KeyPath" type="xsd:string" minOccurs="0"/>
            <xsd:element name="CertificatePath" type="xsd:string" minOccurs="0"/>
            <xsd:element name="ProxyPath" type="xsd:string" minOccurs="0"/>
            <xsd:element name="CACertificatesDir" type="xsd:string" minOccurs="0"/>
            <!-- Retry count. Specifie how many times have an ISIS to timeout to handle it as
                 unavailable. If there is no value specified the default value: 5 will be used. -->
            <xsd:element name="Retry" type="xsd:string" minOccurs="0" maxOccurs="1"/>
        </xsd:sequence>
        <!-- Attribute id may be used to refer to this element -->
        <xsd:attribute name="id" type="xsd:string" use="optional"/>
    </xsd:complexType>
    <xsd:element name="InfoRegistrar" type="iregc:InfoRegistrar_Type"/>

>>>>

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{soul}				       %package for highlight text
%\usepackage{trackchanges}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2CM}
\hypersetup{
  pdfauthor = {},
  pdftitle = {Documentation of the Web Service based ARC Information System},
  pdfsubject = {},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

%%%%\vspace*{1.5cm}
%%%%---- A list of authors ----
%%%%    {\centering \large \footnote{@} \large \par}
\end{titlepage}

@@

\subsection{Functionality}
\label{sub:peer_to_peer_functionality}
The ISIS nodes are able to organize themselves into a peer-to-peer network by synchronizing the stored replicated database and maintaining this synchronicity and the network topology. Every node of this peer-to-peer cloud (that is an ISIS service) has its own PeerID, an identifier only used in this network, made as a hash of its endpoint URL. The members are ordered by this identifier in a ring topology, where the successor entity of the ISIS with the largest PeerID is the ISIS with the smallest PeerID. This metric is used to define the neighbor relations that are the basis of the inter-peer-to-peer network communication. It is presumed that only registered ISIS nodes can have any role in the cloud.

\subsubsection{Connecting to the peer-to-peer network}
Every ISIS tries to connect to the network at its startup by accessing the InfoProvider ISIS nodes known from its configuration. The choice from among the few pre-configured InfoProviders is random to achieve as much load-balancing as possible. (In the design it is assumed that a huge amount of ISIS nodes are configured in very similar way, where the InfoProvider ISIS nodes are mainly the same.) If there is no InfoProvider available or nothing is known from the configuration, the node decides to be the first member and to build a new network.

If there are any InfoProvider nodes available the list of member ISIS nodes will be queried and the `successor' will be defined using the previously mentioned metric. This `successor' node will help the entering ISIS to connect to the cloud. The new member will use the `Connect' Web Service operation to get the newest version of the database stored in the peer-to-peer network, and place its own entries in the cloud by registering them. An initial database synchronization can be achieved using this method.

We can prevent the InfoProvider ISIS nodes from overloading by balancing the operation of database synchronizing --- this being the most costly procedure in the connection phase.

There is also an other extension used regarding the InfoProvider ISIS nodes. If a node is not able to connect to any of the InfoProviders then it tries to repeat it later. This results in that the network is able to fuse two disjoint parts of the network when the missing InfoProvider ISIS appears again.

\subsubsection{Routing in the peer-to-peer network}
We are keeping the databases in synch by passing every Registration and RemoveRegistrations to every node acting in the network. This yields redundancy and fault tolerant message routing between any two ISIS nodes.

The most important parameter of the network is the {\it sparsity}. This is an integer not less than 2. It determines the number of neighbors as a function of the actual number of member nodes of the network, where the number of neighbors can change when new members connect to the network or old members leave. A sparse graph can be shaped by choosing a great value for sparsity or a dense one by choosing a smaller one. This value is defined during the service configuration and determines the main properties of the peer-to-peer network.
\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{Circle2-1.png}}}
\caption{\label{fig:p2p_neighbors_2} The neighbor connections in the peer-to-peer network in case of sparsity=2} }
\end{figure}

@@

The registration of a service is carried out by an internal module called \hl{Registrant}. The \hl{Registrant} is an active module of the HED (Hosting Environment Daemon) which is bound to a set of ISISes. In practice, the configuration part of the \hl{Registrant} contains exactly one ISIS to bind, and the \hl{Registrant} will collect the necessary information about the other ISISes belonging to the same network.

To register services to more than one ISIS network multiple \hl{Registrant} instances has to be configured. In this case, the default \hl{Registrant} will be used for registering every service unless configured explicitly.
The registration of a service can be done once or periodically based either on the configuration of the \hl{Registrant} or overwritten for every service separately. The \hl{Registrant} is also performing message aggregation of all services linked to it if possible. The simplified algorithm of the \hl{Registrant} is presented below.

\begin{framed}
  \hl{Registrant} --- pseudo algorithm\\
  \\
  \verb#// Initialize phase#\\
  \verb#Read the configuration and store the information about the services in a list#\\
  \verb#do { // Cyclic phase in a different Thread#\\
  \verb# wake_up_time = now();#\\
  \verb# messages = null;#\\
  \verb#  if ( 0 < count(service where service.next_run <= wake_up_time)) {#\\
  \verb#    foreach( service where service.next_run <= wake_up_time) {#\\
  \verb#      messages.add(service.RegistrationCollector);#\\
  \verb#      service.next_run = wake_up_time + service.period;#\\
  \verb#    }#\\
  \verb#    if (0 < count(messages)) {#\\
  \verb#      sent_message = assemble message with headers(messages);#\\
  \verb#      send(sent_message);#\\
  \verb#    }#\\
  \verb#  } else {#\\
  \verb#    sleep(min(service.next_run) - now()); #\\
  \verb#  }#\\
  \verb#} while(true)#\\
\end{framed}
The current implementation does not allow the value of the period to be less than 2 minutes.

The service provides the \textbf{Service Advertisement} part of the information sent to ISIS (see Section \ref{sub:isis_interface}).
Before sending this information the \hl{Registrant} extends it with additional data (\textbf{Service Advertisement Metadata}).

An example layout of services is shown on Figure \ref{fig:registration_overview}. In this configuration example 
there are two \hl{Registrant}s configured in one HED container for three services. The \textit{\hl{Registrant} A} is the 
default \hl{Registrant} and the services configured in the following way:

\begin{itemize}
  \item Service 1: There is no \hl{Registrant} configured so the default one will register it.
  \item Service 2: The \textit{\hl{Registrant} A} is configured explicitly.
  \item Service 3: The \textit{\hl{Registrant} B} is configured explicitly.
\end{itemize}

So Service 1 and 2 are both handled by \textit{\hl{Registrant} A} and Service 3 by \textit{\hl{Registrant} B}.
In the first step each \hl{Registrant} performs information collection from all assigned services sequentially. \hl{Registrant}s themselves are executing in parallel. For Figure \ref{fig:registration_overview} the approximate sequence of information collection is (\textit{A1}-\textit{A2}, \textit{B1}-\textit{B2}) followed by (\textit{A3}, \textit{A4}). After collection  the aggregated information is sent to the corresponding ISISes independently by each other \hl{Registrant} (\textit{A5}, \textit{B5}).

@@

For every piece of information provided by the service through the LIDI interface the same procedure as described in 
Section \ref{sub:isis_authorization_information_authorization} should be applied.

@@

The registration operation is done by the \hl{Registrant}s as described in Section \ref{sec:service_registration}.
Each \hl{Registrant} is instantiated by the corresponding InfoRegistrar XML element in the configuration file as 
defined in Section \ref{annex:service_configuration_schema}. The first InfoRegistrar defines the \hl{Registrant} 
that is used to register every service by default.

Registration parameters per service are defined by the InfoRegister element located inside the corresponding 
service configuration element. To assign a service to specific \hl{Registrant}(s) one may use Registrar 
configuration elements situated inside InfoRegister. If no Registrar element is defined the service will 
be registered using first (default) \hl{Registrant}. To make a service not register itself the special configuration element 
NoRegister has to be used.

The Following elements are defined inside the configuration element of the \hl{Registrant}:
\begin{description}
\item{URL} specifies the contact endpoint of a bootstrap ISIS. If needed further ISIS addresses will be 
queried from this service. This element is mandatory.
\item{KeyPath, CertificatePath, ProxyPath and CACertificatesDir} are paths to files storing X509 
credentials used for establishing connections. Those elements are optional and needed only if 
TLS communication is used.
\item{Retry} specifies how many times communication to ISIS have to fail/timeout to start treating it
as unavailable. The default value is 5.
\end{description}

@@

    <xsd:complexType name="InfoRegistrar_Type">
        <!-- This element defines configuration of Information Registration
             active element. This element is located at top level configuration
             document. Each instance if such element instatiates one instance
             of Information \hl{Registrant} active element. -->
        <xsd:sequence>
            <!-- URL specifies contact endpoint of a bootstrap Information
                 Registration service. Further ISIS adresses will be queried
                 from this service. -->
            <xsd:element name="URL" type="xsd:string"/> <!-- maxOccurs="unbounded" -->
            <!-- Optional KeyPath, CertificatePath, ProxyPath and CACertificatesDir are
                 paths to files storing X509 credentials used for establishing connections. -->
            <xsd:element name="KeyPath" type="xsd:string" minOccurs="0"/>
            <xsd:element name="CertificatePath" type="xsd:string" minOccurs="0"/>
            <xsd:element name="ProxyPath" type="xsd:string" minOccurs="0"/>
            <xsd:element name="CACertificatesDir" type="xsd:string" minOccurs="0"/>
            <!-- Retry count. Specifie how many times have an ISIS to timeout to handle it as
                 unavailable. If there is no value specified the default value: 5 will be used. -->
            <xsd:element name="Retry" type="xsd:string" minOccurs="0" maxOccurs="1"/>
        </xsd:sequence>
        <!-- Attribute id may be used to refer to this element -->
        <xsd:attribute name="id" type="xsd:string" use="optional"/>
    </xsd:complexType>
    <xsd:element name="InfoRegistrar" type="iregc:InfoRegistrar_Type"/>
